The Mass Distribution of Stellar-Mass Black Holes

### 2010-10-31 paper.tex
All changes in near-final version.

\author{Andrew Cantrell \and Laura Kreidberg \and Charles Bailyn}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements, \changed{taking into account the number of degrees of
    freedom of each model.}  The mass distribution of the low-mass
  systems is best fit by a power-law, while the distribution of the
  complete sample is best fit by the exponential model.  \changed{This
    difference clearly indicates that the character of the low mass
    X-ray binary systems differs from the higher-mass systems.}  We
  examine the existence of a ``gap'' between the most massive neutron
  stars and least massive black holes, finding that the best model
  (the power law) fitted to the low-mass systems only gives a minimum
  black-hole mass of 4.3 $\Msun$ (90\% confidence), while the best
  model (the exponential) fitted to all 20 systems gives a minumum
  black-hole mass of 4.5 $\Msun$ (90\% confidence).  We conclude that
  our sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
  \changed{Our results on the low-mass sample are in qualitative
    agreement with those of \citet{Ozel2010}, although our broad model
    selection analysis more reliably reveals the best-fit quantitative
    description of the underlying mass distribution.  The results on
    the complete sample of low- and higher-mass systems are in
    qualitative agreement with \citet{Fryer2001} although the presence
    of a mass gap remains theoretically unexplained.}
\end{abstract}

@@

Observations of low-mass X-ray binaries in both the optical and X-ray
can provide a measurement of the mass of the compact object in these
systems.  \changed{The current sample of dynamically measured stellar
  mass black holes includes 15 low-mass, Roche lobe overflowing
  systems and 5 wind-fed higher-mass systems.  Hence, sophisticaed
  statistical analyses are possible.}

The first study of the mass distribution of stellar-mass black holes,
in \citet{Bailyn1998}, examined a sample of seven low-mass X-ray
binaries thought to contain a black hole, concluding in a Bayesian
analysis that the mass function was strongly peaked around seven solar
masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
least massive black hole and a ``safe'' upper limit for neutron star
masses of 3 $\Msun$ (e.g.\ \citet{Kalogera1996}).  Such a gap is
puzzling in light of theoretical studies that predict a continuous
distribution of compact object supernova remnant masses with a smooth
transition from neutron stars to black holes \citep{Fryer2001}.
\changed{(It should be noted that \citet{Fryer2001} considered binary
  evolution effects only heuristically and put forward some possible
  explanations for the gap from \citet{Bailyn1998} both in the context
  of selection effects or in connection to the energetics of supernova
  explosions.)}

\changed{Towards the end of our anlysis work, we became aware of a
  more recent study \citep{Ozel2010}, also in a Bayesian framework,
  analyzing the low-mass X-ray binary sample.}  It largely confirmed
these findings, examining 16 low-mass X-ray binary systems containing
black holes and finding a strongly peaked distribution at $7.8 \pm 1.2
\, \Msun$.  \citet{Ozel2010} use two models for the mass function: a
Gaussian and a decaying exponential with a minimum ``turn-on'' mass
(motivated by the analytical model of the black-hole mass function in
\citet{Fryer2001}).  \changed{\citet{Ozel2010} do not provide
  confidence limits for the minimum black hole mass, instead
  discussing only the model parameters at the peak of their posterior
  distributions.  They also do not perform any model selection
  analysis; thus, they give the distribution of parameters within each
  of their models, but cannot say which model is more likely to
  correspond to the true distribution of black hole masses.}
Nevertheless, it appears that their analysis confirms the existence of
a mass gap.  \citet{Ozel2010} discuss possible selection effects that
could lead to the appearance of a mass gap, but conclude these effects
could not produce the observed gap, which they therefore claim is a
real property of the black hole mass distribution.

@@

We use a Bayesian Markov-Chain Monte Carlo (MCMC) analysis to
quantitatively assess a wide range of models for the black hole mass
function for both samples.  We include both parameteric models, such
as a Gaussian, and non-parameteric models where the mass function is
represented by histograms with various numbers of bins.  (Our set of
models includes those of \citet{Ozel2010} and \citet{Bailyn1998}.)
After computing posterior distributions for the model parameters, we
use model selection techniques (including a new technique for
efficient reversible-jump MCMC \citep{Farr2010}) to compare the
evidence for the various models from both samples.

In agreement with \citet{Ozel2010} and \citet{Bailyn1998}, we find
strong evidence for a mass gap among the best models for both samples,
with the best model for the low-mass systems giving a minimum black
hole mass of 4.3 $\Msun$ (at 90\% confidence), and the best model for
the combined sample of lower- and high-mass systems giving a minimum
black hole mass of 4.5 $\Msun$ (at 90\% confidence).  Further, among
the models with lower evidence, most also have a mass gap.

@@

@@

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|l|}
      \hline
      Source & $f$ ($\Msun$) & $q$ & $i$ (degrees) & References \\
      \hline \hline
      GRS 1915 & $N(9.5, 3.0)$ & $N(0.0857, 0.0284)$ & $N(70, 2)$ &
      \citet{Greiner2001} \\
      XTE J1118 & $N(6.44, 0.08)$ & $N(0.0264, 0.004)$ & $N(68, 2)$ &
      \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
      XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
      \cite{Orosz2004} \\
      GRS 1009 & $N(3.17, 0.12)$ & $N(0.137, 0.015)$ & $I(37, 80)$ &
      \cite{Filippenko1999} \\
      A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
      \citet{Cantrell2010} \\ & & & & \citet{Neilsen2008} \\
      GRO J0422 & $N(1.13, 0.09)$ & $U(0.076, 0.31)$ & $N(45, 2)$ &
      \citet{Gelino2003} \\
      Nova Mus 1991 & $N(3.01, 0.15)$ & $N(0.128, 0.04)$ & $N(54,1.5)$
      & \cite{Gelino2001} \\
      GRO J1655 & $N(2.73,0.09)$ & $N(0.3663, 0.04025)$ & $N(70.2,
      1.9)$ & \citet{Greene2001} \\
      4U 1543 & $N(0.25, 0.01)$ & $U(0.25, 0.31)$ & $N(20.7,1.5)$ & 
      \citet{Orosz2003} \\
      XTE J1550 & $N(7.73,0.4)$ & $U(0,0.04)$ & $N(74.7, 3.8)$ &
      \citet{Orosz2010} \\
      V4641 Sgr & $N(3.13,0.13)$ & $U(0.42,0.45)$ & $N(75,2)$ &
      \citet{Orosz2003} \\
      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $I(66, 70)$ &
      \citet{Charles2006} \\
      & & & & \citet{Khargharia2010} \\
      GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
      \citet{Casares2009} \\
      Nova Oph 77 & $N(4.86,0.13)$ & $U(0, 0.053)$ & $I(60, 80)$ &
      \citet{Charles2006} \\
      GS 2000 & $N(5.01, 0.12)$ & $U(0.035, 0.053)$ & $I(43, 74)$ &
      \citet{Charles2006} \\
      \hline \hline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% High Mass Follows
      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
      \citet{Gies2003} \\ 
      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
      \citet{Orosz2007} \\
      NGC 300 X1 & $N(2.6, 0.3)$ & $U(1.05, 1.65)$ & $I(60, 75)$ &
      \citet{Crowther2010} \\
      LMC X1 & $N(0.148, 0.004)$ & $N(2.91, 0.49)$ & $N(36.38, 2.02)$
      & \citet{Orosz2009} \\
      IC 10 X1 & $N(7.64, 1.26)$ & $U(0.7, 1.7)$ & $I(75, 90)$ & 
      \citet{Prestwich2007} \\
       & & & & \citet{Silverman2008} \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:sources} The source parameters for the 20 low
    mass X-ray binaries used in this work.  The first 15 systems have
    low-mass secondaries that feed the black hole via Roche lobe
    overflow; the last five systems have high-mass secondaries ($q
    \gtrsim 1$) that feed the black hole via winds.  In each line, $f$
    is the mass function for
    the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
    inclination of the system to the line of sight.  We indicate the
    distribution used for the true parameters when computing the
    probability distributions for the masses of these systems:
    $N(\mu,\sigma)$ implies a Gaussian with mean $\mu$ and standard
    deviation $\sigma$, $U(a,b)$ is a uniform distribution between $a$ and
    $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
    angles $\alpha$ and $\beta$.}
\end{table}

@@

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Previous works
\citep{Ozel2010,Bailyn1998} have considered models with only two
parameters; for such models evaluating $p(\vtheta|d)$ on a grid may be
a reliable method.  Many of our models for the underlying mass
distribution have three or more parameters.  Exploring the entirety of
these parameter spaces with a grid rapidly becomes prohibitive as the
number of parameters increases.  A more efficient way to explore the
distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
samples, $\{ \vtheta_i \, | \, i = 1, \ldots \}$, such that a
particular parameter sample, $\vtheta$, appears in the sequence with a
frequency equal to its posterior probability, $p(\vtheta|d)$.  In this
way, regions of parameter space where $p(\vtheta|d)$ is large are
sampled densely while regions where $p(\vtheta|d)$ is small are
effectively ignored.

@@

An important part of any Bayesian analysis is the priors placed on the
parameters of the model.  The choice of priors can bias the results of
the analysis through both the shape and the range of prior support in
parameter space.  The prior should reflect the ``best guess'' for the
distribution of parameters before examining any of the data.  In the
absence of any information about the distribution of parameters, it is
best to choose a prior that is broad and uniformative to avoid biasing
the posterior as much as possible.

@@

\citet{Fryer2001} studied the relation between progenitor and remnant
mass in simulations of supernova explosions.  Combining this with the
mass function for supernova progenitors, they suggested that the
black-hole mass distribution may be well-represented by a decaying
exponential with a minimum mass: \fixme{Vicky, do you want to say
  anything more here?}
\begin{equation}
  \label{eq:exp-def}
  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
    \right] & M \geq \Mmin \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose a prior for this model where $\Mmin$ is uniform between 0
and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
range ensuring that $40\Msun$ is at least two scale masses above the
cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
The resulting prior is
\begin{equation}
  p(\vtheta) = p(\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
    + 2 M_0 \leq 40, \\
    0 & \textnormal{otherwise}
  \end{cases}
\end{equation}

@@

Though we do not expect to find a second peak in the low-mass
distribution, we may find evidence of one when exploring the combined
low- and higher-mass samples.  To look for a second peak in the
black-hole mass distribution, we use a two-Gaussian model:
\begin{multline}
  \label{eq:two-gaussian-def}
  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
\end{multline}
The probability is a linear combination of two Gaussians with weight
$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
probability lies below $40 \Msun$ with the prior 
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
    \mu_2, \quad 0 \leq \alpha \leq 1 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
in Equation \eqref{eq:gaussian-prior-def}.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distributions for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian is centered around the peaks of the low-mass
distributions.  The second Gaussian's parameter distributions are much
broader.  The second Gaussian appears to be sampling the tail of the
mass samples.  In spite of the ability to match a more complicated
distribution, we find that this model is strongly disfavored relative
to the single-Gaussian model for this dataset:
$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
4.7$ (see Sections \ref{sec:model-selection} and
\ref{sec:low-mass-model-selection} for discussion).

@@

\subsubsection{Low-Mass Model Selection}
\label{sec:low-mass-model-selection}
We have performed a suite of 500 independent reversible-jump MCMCs
jumping between all the models (both parametric and non-parametric)
described in Section \ref{sec:models} using the single-model MCMC
samples to construct an efficient jump proposal for each model as
described above (see Appendix \ref{sec:reversible-jump-mcmc}).  The
numbers of counts in each model are consistent across the MCMCs in the
suite; Figure \ref{fig:rj} displays the average probability for each
model across the suite, along with the 1-$\sigma$ errors on the
average inferred from the standard deviation of the model counts
across the suite.  Table \ref{tab:rj} gives the numerical values of
the average probability for each model across the suite of MCMCs.

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} The relative probability of the models
    discussed in Section \ref{sec:models} as computed using the
    reversible-jump MCMC with the efficient jump proposal algorithm
    described in Section \ref{sec:reversible-jump-mcmc}.  In
    increasing order along the $x$-axis, the models are the power-law
    of Section \ref{sec:power-law} (PL), the decaying exponential of
    Section \ref{sec:exponential} (E), the single Gaussian of Section
    \ref{sec:gaussian} (G), the double Gaussian of Section
    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
    five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The average of 500 independent reversible-jump
    MCMCs is plotted, along with the 1-$\sigma$ error on the average
    inferred from the standard deviation of the probability from the
    individual MCMCs.  As discussed in the text, the power-law and
    Gaussian models are the most favored.}
\end{figure}

@@

\subsubsection{High-Mass Model Selection}
\label{sec:high-mass-model-selection}

>>>>

\author{Andrew Cantrell \and Laura Kreidberg \and Charles D. Bailyn}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements, \changed{taking into account the number of degrees of
    freedom of each model.}  The mass distribution of the low-mass
  systems is best fit by a power-law, while the distribution of the
  complete sample is best fit by the exponential model.  \changed{This
    difference clearly indicates that the character of the low mass
    X-ray binary systems differs from the higher-mass systems.}  We
  examine the existence of a ``gap'' between the most massive neutron
  stars and least massive black holes.  \changed{Our analysis
    generates distributions of lower bounds on the black hole mass
    distribution; we use the 1\% quantile of the mass distribution,
    $M_{1\%}$, for a lower bound because some of our distributions do
    not have a sharp lower limit.  For the best model (the power law)
    fitted to the low-mass systems this distribution of lower-bounds
    has $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence, while the best
    model (the exponential) fitted to all 20 systems has $M_{1\%} >
    4.5$ $\Msun$ with 90\% confidence.}  We conclude that our sample
  of black hole masses provides strong evidence of a gap between the
  maximum neutron star mass and \changed{the lower bound on} black
  hole masses.  \changed{Our results on the low-mass sample are in
    qualitative agreement with those of \citet{Ozel2010}, although our
    broad model selection analysis more reliably reveals the best-fit
    quantitative description of the underlying mass distribution.  The
    results on the complete sample of low- and higher-mass systems are
    in qualitative agreement with \citet{Fryer2001} although the
    presence of a mass gap remains theoretically unexplained.}
\end{abstract}

@@

Observations of low-mass X-ray binaries in both the optical and X-ray
can provide a measurement of the mass of the compact object in these
systems.  \changed{The current sample of stellar mass black holes with
  dynamically measured masses includes 15 low-mass, Roche lobe
  overflowing systems and 5 wind-fed higher-mass systems.  Hence,
  sophisticated statistical analyses of the black hole mass
  distribution in these systems are possible.}

The first study of the mass distribution of stellar-mass black holes,
in \citet{Bailyn1998}, examined a sample of seven low-mass X-ray
binaries thought to contain a black hole, concluding in a Bayesian
analysis that the mass function was strongly peaked around seven solar
masses.  \changed{(A similar analysis of the neutron star mass
  distribution can be found in \citet{Finn1994}.)}  \citet{Bailyn1998}
found evidence of a ``gap'' between the least massive black hole and a
``safe'' upper limit for neutron star masses of 3 $\Msun$ (e.g.\
\citet{Kalogera1996}).  Such a gap is puzzling in light of theoretical
studies that predict a continuous distribution of compact object
supernova remnant masses with a smooth transition from neutron stars
to black holes \citep{Fryer2001}.  \changed{(But note that
  \citet{Fryer2001} considered binary evolution effects only
  heuristically and put forward some possible explanations for the gap
  from \citet{Bailyn1998} both in the context of selection effects or
  in connection to the energetics of supernova explosions.)}

\changed{Towards the end of our analysis work, we became aware of a
  more recent study \citep{Ozel2010}, also in a Bayesian framework,
  analyzing the low-mass X-ray binary sample.}  It largely confirmed
these findings, examining 16 low-mass X-ray binary systems containing
black holes and finding a strongly peaked distribution at $7.8 \pm 1.2
\, \Msun$.  \citet{Ozel2010} use two models for the mass function: a
Gaussian and a decaying exponential with a minimum ``turn-on'' mass
(motivated by the analytical model of the black-hole mass function in
\citet{Fryer2001}).  \changed{\citet{Ozel2010} do not provide
  confidence limits for the minimum black hole mass, instead
  discussing only the model parameters at the peak of their posterior
  distributions.  They also do not perform any model selection
  analysis; thus, they give the distribution of parameters within each
  of their models, but cannot say which model is more likely to
  correspond to the true distribution of black hole masses.}
Nevertheless, it appears that their analysis confirms the existence of
a mass gap.  \citet{Ozel2010} discuss possible selection effects that
could lead to the appearance of a mass gap, but conclude these effects
could not produce the observed gap, which they therefore claim is a
real property of the black hole mass distribution.

@@

We use a Bayesian Markov-Chain Monte Carlo (MCMC) analysis to
quantitatively assess a wide range of models for the black hole mass
function for both samples.  We include both parametric models, such
as a Gaussian, and non-parametric models where the mass function is
represented by histograms with various numbers of bins.  (Our set of
models includes those of \citet{Ozel2010} and \citet{Bailyn1998}.)
After computing posterior distributions for the model parameters, we
use model selection techniques (including a new technique for
efficient reversible-jump MCMC \citep{Farr2010}) to compare the
evidence for the various models from both samples.

\changed{We define the ``minimum black hole mass'' to be the 1\%
  quantile, $M_{1\%}$, in the black hole mass distribution (see
  Section \ref{sec:minimum-mass}).  In agreement with \citet{Ozel2010}
  and \citet{Bailyn1998}, we find strong evidence for a mass gap among
  the best models for both samples.  Our analysis gives distributions
  for $M_{1\%}$ implied by the data in the context of each of our
  models for the black hole mass distribution.  In the context of the
  best model for the low-mass systems (a power-law), the distribution
  for $M_{1\%}$ has $M_{1\%} > 4.3$ $\Msun$ (at 90\% confidence); in
  the context of the best model for the combined sample of lower- and
  high-mass systems the distribution of $M_{1\%}$ has $M_{1\%} > 4.5$
  $\Msun$ with 90\% confidence.  Further, in the context of models
  with lower evidence, most also have a mass gap, with 90\% confidence
  bounds on $M_{1\%}$ significantly above a ``safe'' maximum neutron
  star mass of 3 $\Msun$ \citep{Kalogera1996}.}

@@

\changed{The structure of this paper is as follows.  In Section
  \ref{sec:systems} we discuss the 15 systems that comprise the
  low-mass X-ray binary black hole sample and the 5 additional
  higher-mass, wind-fed systems that make up the complete sample.  In
  Section \ref{sec:models} we discuss the Bayesian techniques we use
  to analyze the black hole mass distribution, the techniques we use
  for model selection, and the parametric and non-parametric models we
  will for the black hole mass distribution.  In Section
  \ref{sec:results} we discuss the results of our analysis and model
  selection.  In Section \ref{sec:minimum-mass} we discuss the
  distribution of the minimum black hole mass implied by the analysis
  of Section \ref{sec:results}.  In Section \ref{sec:conclusion} we
  summarize our results and comment on the significance of the
  observed mass gap in the context of theoretical models.  Appendix
  \ref{sec:mcmc} describes MCMC techniques in some detail.  Appendix
  \ref{sec:reversible-jump-mcmc} explains our novel algorithm for
  efficiently performing the reversible jump MCMC computations used in
  the model comparisons of Section \ref{sec:results} (but see also
  \citet{Farr2010}).}

@@

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|l|}
      \hline
      Source & $f$ ($\Msun$) & $q$ & $i$ (degrees) & References \\
      \hline \hline
      GRS 1915 & $N(9.5, 3.0)$ & $N(0.0857, 0.0284)$ & $N(70, 2)$ &
      \citet{Greiner2001} \\
      XTE J1118 & $N(6.44, 0.08)$ & $N(0.0264, 0.004)$ & $N(68, 2)$ &
      \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
      XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
      \cite{Orosz2004} \\
      GRS 1009 & $N(3.17, 0.12)$ & $N(0.137, 0.015)$ & $I(37, 80)$ &
      \cite{Filippenko1999} \\
      A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
      \citet{Cantrell2010} \\ & & & & \citet{Neilsen2008} \\
      GRO J0422 & $N(1.13, 0.09)$ & $U(0.076, 0.31)$ & $N(45, 2)$ &
      \citet{Gelino2003} \\
      Nova Mus 1991 & $N(3.01, 0.15)$ & $N(0.128, 0.04)$ & $N(54,1.5)$
      & \cite{Gelino2001} \\
      GRO J1655 & $N(2.73,0.09)$ & $N(0.3663, 0.04025)$ & $N(70.2,
      1.9)$ & \citet{Greene2001} \\
      4U 1543 & $N(0.25, 0.01)$ & $U(0.25, 0.31)$ & $N(20.7,1.5)$ & 
      \citet{Orosz2003} \\
      XTE J1550 & $N(7.73,0.4)$ & $U(0,0.04)$ & $N(74.7, 3.8)$ &
      \citet{Orosz2010} \\
      V4641 Sgr & $N(3.13,0.13)$ & $U(0.42,0.45)$ & $N(75,2)$ &
      \citet{Orosz2003} \\
      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $I(66, 70)$ &
      \citet{Charles2006} \\
      & & & & \citet{Khargharia2010} \\
      GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
      \citet{Casares2009} \\
      Nova Oph 77 & $N(4.86,0.13)$ & $U(0, 0.053)$ & $I(60, 80)$ &
      \citet{Charles2006} \\
      GS 2000 & $N(5.01, 0.12)$ & $U(0.035, 0.053)$ & $I(43, 74)$ &
      \citet{Charles2006} \\
      \hline \hline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% High Mass Follows
      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
      \citet{Gies2003} \\ 
      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
      \citet{Orosz2007} \\
      NGC 300 X1 & $N(2.6, 0.3)$ & $U(1.05, 1.65)$ & $I(60, 75)$ &
      \citet{Crowther2010} \\
      LMC X1 & $N(0.148, 0.004)$ & $N(2.91, 0.49)$ & $N(36.38, 2.02)$
      & \citet{Orosz2009} \\
      IC 10 X1 & $N(7.64, 1.26)$ & $U(0.7, 1.7)$ & $I(75, 90)$ & 
      \citet{Prestwich2007} \\
       & & & & \citet{Silverman2008} \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:sources} The source parameters for the 20 X-ray binaries used in this work.  The first 15 systems have
    low-mass secondaries that feed the black hole via Roche lobe
    overflow; the last five systems have high-mass secondaries ($q
    \gtrsim 1$) that feed the black hole via winds.  In each line, $f$
    is the mass function for
    the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
    inclination of the system to the line of sight.  We indicate the
    distribution used for the true parameters when computing the
    probability distributions for the masses of these systems:
    $N(\mu,\sigma)$ implies a Gaussian with mean $\mu$ and standard
    deviation $\sigma$, $U(a,b)$ is a uniform distribution between $a$ and
    $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
    angles $\alpha$ and $\beta$.}
\end{table}

@@

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Previous works
\citep{Ozel2010,Bailyn1998} have considered models with only two
parameters; for such models evaluating $p(\vtheta|d)$ on a grid may be
a reliable method.  Many of our models for the underlying mass
distribution have three or more parameters.  Exploring the entirety of
these parameter spaces with a grid rapidly becomes prohibitive as the
number of parameters increases.  A more efficient way to explore the
distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
(MCMC) method (see Appendix \ref{sec:mcmc}).  MCMC methods produce a
chain (sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1,
\ldots \}$, such that a particular parameter sample, $\vtheta$,
appears in the sequence with a frequency equal to its posterior
probability, $p(\vtheta|d)$.  In this way, regions of parameter space
where $p(\vtheta|d)$ is large are sampled densely while regions where
$p(\vtheta|d)$ is small are effectively ignored.

@@

An important part of any Bayesian analysis is the priors placed on the
parameters of the model.  The choice of priors can bias the results of
the analysis through both the shape and the range of prior support in
parameter space.  The prior should reflect the ``best guess'' for the
distribution of parameters before examining any of the data.  In the
absence of any information about the distribution of parameters, it is
best to choose a prior that is broad and uninformative to avoid biasing
the posterior as much as possible.

@@

\citet{Fryer2001} studied the relation between progenitor and remnant
mass in simulations of supernova explosions.  Combining this with the
mass function for supernova progenitors, they suggested that the
black-hole mass distribution may be well-represented by a decaying
exponential with a minimum mass: 
\begin{equation}
  \label{eq:exp-def}
  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
    \right] & M \geq \Mmin \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose a prior for this model where $\Mmin$ is uniform between 0
and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
range ensuring that $40\Msun$ is at least two scale masses above the
cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
The resulting prior is
\begin{equation}
  p(\vtheta) = p(\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
    + 2 M_0 \leq 40, \\
    0 & \textnormal{otherwise}
  \end{cases}
\end{equation}

@@

Though we do not expect to find a second peak in the low-mass
distribution, we may find evidence of one when exploring the combined
low- and higher-mass samples.  To look for a second peak in the
black-hole mass distribution, we use a two-Gaussian model:
\begin{multline}
  \label{eq:two-gaussian-def}
  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
\end{multline}
The probability is a linear combination of two Gaussians with weights
$\alpha$ and $1-\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose
combined conditions on $\mu_i$ and $\sigma_i$ that ensure that most of
the mass probability lies below $40 \Msun$ with the prior
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
    \mu_2, \quad 0 \leq \alpha \leq 1 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
in Equation \eqref{eq:gaussian-prior-def}.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the Gaussian with
smaller mean.  The distributions for $\mu_1$ and $\sigma_1$ are
similar to those of the single Gaussian displayed in Figure
\ref{fig:gaussian}, indicating that this Gaussian is centered around
the peaks of the low-mass distributions.  The second Gaussian's
parameter distributions are much broader.  The second Gaussian appears
to be sampling the tail of the mass samples.  In spite of the extra
degrees of freedom in this model, we find that this model is strongly
disfavored relative to the single-Gaussian model for this dataset:
$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
4.7$ (see Sections \ref{sec:model-selection} and
\ref{sec:low-mass-model-selection} for discussion).

@@

\subsubsection{Model Selection for the Low-Mass Sample}
\label{sec:low-mass-model-selection}
We have performed a suite of 500 independent reversible-jump MCMCs
jumping between all the models (both parametric and non-parametric)
described in Section \ref{sec:models} using the single-model MCMC
samples to construct an efficient jump proposal for each model as
described above (see Appendix \ref{sec:reversible-jump-mcmc}).  The
numbers of counts in each model are consistent across the MCMCs in the
suite; Figure \ref{fig:rj} displays the average probability for each
model across the suite, along with the 1-$\sigma$ errors on the
average inferred from the standard deviation of the model counts
across the suite.  Table \ref{tab:rj} gives the numerical values of
the average probability for each model across the suite of MCMCs.

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} The relative probability of the models
    discussed in Section \ref{sec:models} as computed using the
    reversible-jump MCMC with the efficient jump proposal algorithm
    described in Section \ref{sec:reversible-jump-mcmc}.  (See also
    Table \ref{tab:rj}.)  In increasing order along the $x$-axis, the
    models are the power-law of Section \ref{sec:power-law} (PL), the
    decaying exponential of Section \ref{sec:exponential} (E), the
    single Gaussian of Section \ref{sec:gaussian} (G), the double
    Gaussian of Section \ref{sec:gaussian} (TG), and the one-, two-,
    three-, four-, and five-bin histogram models of Section
    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
    respectively).  The average of 500 independent reversible-jump
    MCMCs is plotted, along with the 1-$\sigma$ error on the average
    inferred from the standard deviation of the probability from the
    individual MCMCs.  As discussed in the text, the power-law and
    Gaussian models are the most favored.}
\end{figure}

@@

\subsubsection{Model Selection for the Complete Sample}
\label{sec:high-mass-model-selection}
